---
title: "Dataframes - In Class lab"
author: "Dr. Shoemaker"
date: "DATA 2401"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Working with built-in data sets

Load R's "USPersonalExpenditure" dataset using the `data()` function

This will produce a data frame called `USPersonalExpenditure`
```{r}
data("USPersonalExpenditure")
```

The variable `USPersonalExpenditure` is now accessible to you. Unfortunately, it's not a data frame (it's actually a matrix, you can find that out using) 

Test this using the `is.data.frame()` function:

```{r}
is.data.frame(USPersonalExpenditure)
```

Luckily, you can pass the USPersonalExpenditure variable as an argument to the `data.frame()` function to convert it a data frame.

Do this, storing the result in a new variable

```{r}
US_spend <- data.frame(USPersonalExpenditure)
```


What are the current column names of your dataframe?
```{r}
colnames(US_spend)
```


Consider: why are they so strange? (Think about whether you could use a number like 1940 with dollar notation!)

What are the current row names of your dataframe?
```{r}
rownames(USPersonalExpenditure)

```


Add a column "Category" to your data frame that contains the rownames

```{r}
USA_spend&Catgeory <- 

```

Consider how this data isn't tidy. Use the `pivot_longer()` function from last class to make the data tidy. 

```{r}

```


How much money was spent on personal care in 1940?
```{r}

```


How much money was spent on Food and Tobacco in 1960?
```{r}


```


What was the highest expenditure category in 1960?
```{r}


```


Define a function `lowest_category` that takes in a year as a parameter, and returns the lowest spending category of that year
```{r}



```

```{r}

```


Using your function, determine the lowest spending category of each year
Hint: use the `sapply()` function to apply your function to a vector of years

```{r}
Vietnam_data <- filter(gapminder, country == "Vietnam", pop >= 1000000, year > 1990 & lifeExp < 80)

Vietnam_data

Vietnam_old_people <- filter(gapminder, country == "Vietnam", year > 1800 & lifeExp < 60)

Vietnam_old_people %>% arrange(desc(year), desc(country))
Vietnam_data %>% filter(year > 1900) %>% arrange(desc(lifeExp))
# order of the control-shift-m shortcut for pipe 
# order of the %>% matter it comes after a function 

data <- read.csv("../")
```
